A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-11 tex/codatadp.tex
my proof

\begin{proof}
  * if two (either directly or indirectly) corecursive constructors
    $\const{C}$ and $\const{D}$, it is
    possible to encode infinitely many alternation patterns,
    e.g. $\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$;
    hence no finite models

  * if one corecursive and one noncorecursive, then possibility to create
    objects of arbitrary depths, e.g. $\const{C}(\ldots(\const{C}(\const{D}))\ldots)$,
    and hence no finite models

  * leaves us with the case of a single corecursive constructor

  * if any of the noncorecursive arguments has cardinality greater than one,
    then it is possible to encode alternation patterns using it,
    e.g. $\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$.

  * otherwise, by the coinduction principle, there is only one value that can
    exist

%  * assume for simplicity no indirect recursion, but this does not radically
%    change the argument
%  * assume there are at least two values built with C.
%    they must be different at some point, e.g.
%    C(0, C(1, C(0, ...))) and C(0, C(1, C(1, ...)))
%    can use that to create infinitely alternation patterns
%  * leaves us with the case of a single constructor
% * since we are looking at a specific model, we can assume all ordinary
%   types are finite datatypes with nullary constructors corresponding to their
%   elements
% * if two
% * either there is only one ``path'' from ctr to itself
%   or at least two;
\qed
\end{proof}

>>>>

\begin{proof}
By definition, the type is equipped with at least one (directly or indirectly)
corecursive constructor $\const{C}$. If it additional has second
corecursive constructor $\const{D}$, it is possible to encode infinitely many
alternation patterns---e.g.,
$\const{C}(\const{D}(\const{C}(\const{C}(\ldots))))$---all of which correspond
to distinct values. If the type has a noncorecursive constructor
$\const{E}$, it is possible to create terms of arbitrary depths---e.g.,
$\const{C}(\ldots(\const{C}(\const{E}))\ldots)$. In both cases, there can be
no finite models.

Therefore, $\const{C}$ must be the only constructor.
If any of its noncorecursive arguments has a cardinality greater than 1,
it is possible to encode alternation patterns using it---e.g.,
$\const{C}(0,\: \const{C}(1,\: \const{C}(0,\: \const{C}(0,\: \ldots))))$---which
again excludes finite models. Otherwise, the coinduction principle ensures
that there can be only one value.
\qed
\end{proof}

%  * assume for simplicity no indirect recursion, but this does not radically
%    change the argument
%  * assume there are at least two values built with C.
%    they must be different at some point, e.g.
%    C(0, C(1, C(0, ...))) and C(0, C(1, C(1, ...)))
%    can use that to create infinitely alternation patterns
%  * leaves us with the case of a single constructor
% * since we are looking at a specific model, we can assume all ordinary
%   types are finite datatypes with nullary constructors corresponding to their
%   elements
% * if two
% * either there is only one ``path'' from ctr to itself
%   or at least two;
